---
title: "STOR 455 Class 3 R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r message=FALSE, warning=FALSE}
# message=FALSE, warning=FALSE supress warnings and messages from appearing in knitted html

library(readr)

DistanceHome <- read_csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/DistanceHome.csv")

# If notebook and csv file are saved in the same folder
# DistanceHome <- read_csv("DistanceHome.csv")

Domestic=subset(DistanceHome,Distance<250)
```
#### Class Notes #### 
# Using 1 quant var to predict another; does a line fit that data? 
# In practice, things aren't going to look super well; we also nee to see "Does a line not fit it well and it's still okay to use the linear process?" 
# Simple Linear Model 
## Y = B0 + B1X + E ( Response = INtercept + Slope(predictor) + Error) 
# Scatter plot in R; 
# Note: Clear your environment frequently; you can use code or you can use the broom button in the global enervionment tab 
# CNTRL + ALT + I = chunck 
```{r}
# Want to do X - Y Axis 
plot(Distance~Hours, Domestic) # Cna doa lot of things; depends on teh data we give it; we want to give it VERTICAL AXIS ~ HORIZONTAL AXIS, Dataset # IF you dont want it to show up below you, then you can not do chunks 
```
#### How does this line best fit the dat? #### 
# Use function lm = linear model; uses the same argumet as plot 
lm(Distance~Hours, data = Domestic) # WIll need to highlight this to read it because apparently if you mix chunks and not chunks, it only want to lok at chunks

mod1 <- lm(Distance~Hours, data = Domestic) #ab line, then will overlap plots 
plot(Distance~Hours, data = Domestic) # Makes the scatter plot 
abline(mod1) #Adds the line to the scatter plot 

#### HOw is this line being made? #### 
# ITs trying tooptimize 
# IT's minimizing something; minimizing the sum of the squared residuals 
# https://www.geogebra.org/m/UxJQorBl
## All the points are draggable; S and T are also dragable; it will change teh intercept if S changes; change T = change in slope 
## Can use this to look at different ways to draw a line of best fit 
## Trying to find a line tha tgets a very low sum of squared residuals 
## Show squared residuals shows teh difference from teh sine to the point; if we drag the line around we can try and find the "best" line; FInds where those sums are being minimized 
## Line of best fit will show wehre the sqaure fo resdiuals is the smallest

#### Simple LInear MOdel COnditions #### 
# We have a working model; but there are a lot of conditions that have to be met to know if we should use this data 
# WHen we make a regression line 
# Check: 
## Linearirty; when you plot data, does it look like a line fits?  Or would soemthing else look better? 
###OTher ways to look at Linearity: Plot the residuals vs the prediction 
#### Look at the residual plot 
#####mod1 <- lm(Distance~Hours, data = DOmestiC) ; then summary stat your way to freedome 
## Zero Mean; THe distribution of errors is centered at zero; we can't check there because of teh least squares regession line - we will always get zero; if you did it by hand you could get roughly zero 
# Constant Variablce; Are theere any places they are tightly packed? Look for a fan-shaped pattern; based on teh residual standard area; look at original scatter plot to see if it's weird 
# INdepdnece; look at the plot with fitted line; does it look similarly packed from L to R; or are there patterns? WE want similiarity; in the data colection, was the data colelcted in a way that may have skewed how the data was collected; to a certain point we can't really know if its true or not; ;it is more dependent on what yo're doing and working with 
# NOrmailty ; WE have all these errors, are they normally distributed; does teh random have a pattern or is it normally distributed ranomness; our current model here doesn't meet this super well 

#### Summary Linear stats #### 
summary(mod1) # Gives you an overview 
# Chunks are useful if you have a large set of code and only want to run one part of it; the ### Divisions help show sections of a whole that need to runa ll togtehr 
mod1$residuals
mod1$fitted.values 
# Look at the above because we want to make a pplot against the fitted values; to see difference 
plot(mod1$residuals~mod1$fitted.values)

# A hor line at 0 = regression line; the vertical distance = teh vertal distance when it had a different slope 
# HOw to put a line into teh groaph 
abline(0,0) # Intercept of 0, and slope of 0) ONly layers a line on the previously run plot 

# Does this horizonal line describe the data? 
# Useful because it shows residual and fitted; it will show all teh predictors; even if you havem ultiple values to look at; 
#### THE ABOVE IS GOOD ALL AROUND #### 

#### How to test if residuals are normal #### 
# Does teh historgram look like a bellcurve shape or not? 
# It's useful if it's NOT a bell curve shape 
# use teh hist finction 
hist(mod1$residuals)
# Shows that we defintiely have a skew; the two points below a line are wehre we have teh neg residuals 
## ALSO Could use a NORMAL QUANTILE PLOT
qqnorm(mod1$residuals) # taes all residuals and basically making them zscores and ordering them from smallest to biggest 
qqline(mod1$residuals) # Adds a line to the qqnorm scatter plot ; if it's noraml, then the scatter plot should fit the line perfectly.  Some values are perfect and otehrs aren't 

#### Randomization and other ####
#Sets randomization
set.seed(455001) # THis will give us the same random number table 

#Sample of 99 values from a Normal distribution; mean=0; sd=1; sorted ascending
x = sort(rnorm(99,0,1)) # Takes 99 values from a norm dist that has a mean of 0 and a std of 1; from a normal dis and how non norm would it be by chance? 
hist(x)

#list of integers 1 through 99.
y = c(1:99) # Array of 1 - 99; 

#z-scores of dataset of 99 values if perfectly normally distributed
normy = qnorm(y/100) # take the noraml distrubtion and see what the area under teh curve is doing for different values; what should be teh vvalue to distribute to separate the lowest 1/100th, then 2/100th, etc. 
# For visual -> google statkey -> normal -> gives normal curve with mean 0, std of 1; then -> Left tail button > will give propostion of area under tail curve > Change that to 1/100th > should have a zscore (how many std the thing is below or aove teh mean) ; if it's perfectly normal then it would be -2.054 for a 2/100th; qq line is telling you what would a perfect dirsticbution look like and are tehre differences with your current set

plot(x~normy)
abline(0,1) # Based on this randomization, the line fits pretty well

#### Generate random values that look like our data #### 
# Summary of teh data we chose; shows that the residual stad error = 18.26
#rnom = takes teh sample of randomly sampled values and we will take 54, which is teh same sides as teh domestic dataset; then use the residual stad error 

x <- rnorm(54,0,18.26)
qqnorm(x)
qqline(x) #Shows you some variability that is okay; how much is okay and how much is not okay? When do those tails fly out? THat's telling you a skew or compact.

#### Homework NOtes #### 
# Sakai has teh pdf file 
# WE can do 1, 2, and 6 based on our current class things 

