---
title: "STOR 455 Class 6 R Notebook"
output:
  html_document:
    df_print: paged
---
 
 rm(list=ls())
```{r message=FALSE, warning=FALSE}
# message=FALSE, warning=FALSE suppress warnings and messages from appearing in knitted files

library(readr)
library(Stat2Data)
```
**Looking at Outliers in Data** 
- Are they real data or not ? DO we want to make a prediction that limits outliers? 
- We still have to worry aout the simple linear model conditions 
- OUtliers can have abig infulence on the SLM Conditions (Simple LInear MOdel (SLM) )
- Unusal points 
  - Are these outlers or infliential points/ 
  - OUtlier = doesn't always have influenc eon a model 
  - OUtlier: a data point that is far from the regression line 
  - influential point" A data p oint that has a large effect on the thing 
  
**Detecting uNusual Cases Overview** 
1. COmputer residueals 
  - raw, standarized, studentized
    - Raw residual - look at it in context; how big is actually big?  
2. Plots of residuals 
  - boxplot and what we've been lookinga t recently 
3. LEverage 
  - next class
4. Cooks distance 
  - next class 
  
```{r}
data("LongJumpOlympics2016")
plot(Gold~Year, jump)
jump1 <- lm(Gold~year, jump)
abline(jump1)
  
# Looking at the ditted residuals; there is acurve, so linearirty migh tbe an issue 
# Code isn't working, but we will still take notes 
# Looks like with teh lm form; every year, increases certain times 
```
```{r}
data("LongJumpOlympics2016")
boxplot(LongJumpOlympics2016$residuals)
# THis isn't working; so we're just going to pay attention to the lecture 
```

```{r}
```

Stand resiudals = like zscore for residuals 
rstandard function = tells you what the standardized thing is 

Very lost already, he's not slowing down 

if you look at the max of gold model? you get the max 
if you want which.max it will tell you where teh value is 

lok at atht filled notebook because struggle buss and code no worky 