---
title: "STOR 455 Class 5 R Notebook"
output:
  html_document:
    df_print: paged
---
rm(list=ls())
# Class Notes
SImple linear model
- Trying to fit the line based onteh sample that we have 
- zero mean is always met with the least squares regression 
- we had to look at: 
- linearliarty, costastnt variance, and normalitiy; contastant variace = how is it stretched out 

# What do do when regression assumptions are violated 
- non line patterns in residuals, etc, 
- DO DATA TRANSFORMATINOS; like nat logs to make the thing look more like what we want 

```{r message=FALSE, warning=FALSE}
# message=FALSE, warning=FALSE suppress warnings and messages from appearing in knitted files

library(readr)
library(Stat2Data)
```
# Notes 
The state2data gives us stuff from our textbook 
```{r}
data("SpeciesArea")
head(SpeciesArea)

# Trying to predict number of speciies based on size of island 
plot(Species~Area, SpeciesArea)
# THis looks like a log graph; a linear pattern isn't going to work 
# DOuble checking tha tit isn't going tow ork 
mod1 <- lm(Species~Area, SpeciesArea)
abline(mod1) # This still looks bad 

plot(mod1, 1:2) #Gives us fitted and QQ Plot 
# Constant variance isn't an issue 
# NOrm of residuals, the tails are a little bad, but we can work with taht 
# Fitted plot, we see what is wrong in teh scatter plot first, its just around the hor line 
```
# What if want to look at outliers and those furthest away? 
HOw far above the line is tha tpoint; we want to use teh max function 
then you can say "which.max" function which tells you which one 
```{r}
max(mod1$residuals)
which.max(mod1$residuals) # Tells us that on row three 
SpeciesArea[3,]
```
# We Can do a transformation that will change it; it's not always going to be a log transformation, but we are going to use it this time 
```{r}
plot(log(Species)~log(Area), SpeciesArea)
mod2 <- lm(log(Species)~log(Area), SpeciesArea)
abline(mod2) # THis looks much more linear 
```
Lets look a little deeper 
```{r}
summary(mod2)

# Pull out things to auto calc predictions
B0 <- summary(mod2)$coef[1,1] # Intercept; want to pull out the coeff table; this is after teh $; the names aren't part of the table so [1,1] will take teh intercept from the lm sum 
B1 <- summary(mod2)$coef[2,1] # Slope; same idea as a bove, but the second row, 1st column

plot(Species~Area, SpeciesArea)
# WE need to solve for the response; need tos olve for species; just doing math to solve for x because we have a function that gives us predicted area 
# exp(B0)^x^B1; we want to plot this 
curve(exp(B0)^x^B1, add = TRUE ) # I want to put this ontop of the plot that I just made, so add "Add = TRUE")

```

# Example 2: Artice Sea Ice #
- Look at teh actual area, the outer boundary of the ice, not account for gaps inside of the area 

```{r}
data("SeaIce")
head(SeaIce)

plot(Extent~t, SeaIce)
modi <- lm(Extent~t, SeaIce)
abline(modi) # THe line is alright 

plot(modi, 1:2)
```

```{r}
plot(modi$residuals~modi$fitted.values)
abline(0,0)

abline(v=c(5.65,6.52), 
       col=c("red", "red"), 
       lty=c(2,2), 
       lwd=c(1, 1)) #Draws red dashed vertical lines; # lwd = width 

plot(modi, 1:2)
```

## Try power transformations ## 
What would happen if raised toa  power? 
HOw does a power change your plot?

```{r}
data("SeaIce")
head(SeaIce)

plot(Extent~t^5, SeaIce)
modi <- lm(Extent^5~t, SeaIce) # If you try raising the X value, R thinks that you're doing it wrong; you need tomake R look at it how you want; need to use insulate I()

plot(modi, 1:2)
```

HOw to plot your transform on the oring data 
```{r}
plot(Extent~t, SeaIce)
B0.i <- summary(modi$coef[1,1])
B1.i <- summary(modi$coef[2,1])

curve((B0.i+B1.i^x)^0.2, add = TRUE) # Doesn't call it, but that's fine, check the notebook 

```