---
title: "STOR 455 Class 7 R Notebook"
output:
  html_document:
    df_print: paged
---

```{r warning=FALSE, message=FALSE}
library(Stat2Data)

data("LongJumpOlympics2016")
data("PalmBeach")
```

**Unusual Points** 
Detecting Overview: 
1. COmpute Residuals (raw, standardized or studentized; if we took that point out of model how does that effect the residuals - we can then see if that point had a lot of influence on the variability the studentized value would be different than teh standardized value)
2. Plot of resiudals, or std. residuals (box plot, scatterplot, norm plot)
3. Leverage 
4. COok's distance 

```{r}
mod1 <- lm(Gold~Year, LongJumpOlympics2016)

plot(Gold~Year, LongJumpOlympics2016)
abline(mod1)

points(LongJumpOlympics2016$Year[16], 
       LongJumpOlympics2016$Gold[16], 
       col = 'red', 
       pch = 16) #16 = the year we want to see the dude; the y = the distance that we want to look at; # pch = the character type; its if you know the list of types of stuff 
# Think abotu how this would look after you knit it 
```
Bob is an outlier; but how much of an ouliter is he? 
*Standarized Resiudals* 
For residuals; mean = 0 and std = about std-hat of the error 
THe standardized resiudals are : std.res = (y-y(hat))/(std and sqrt(leverage))
CHeck the readings for these 

```{r}

summary(mod1)

st.resid.est <- mod1$residuals/summary(mod1)$sigma #WHat is Sigma? 
st.resid.est

rstandard(mod1) - st.resid.est #Leverage 
```

INfluence Analysis 
- Show how if you move a point in the X or Y direction how it impacts teh regression line 

Typical Leverage 
- For a simialr linear model : hi = (1/n) + ((xi-x(bar))^2)/(sum(xi-X(Bar)^2)) <- leverage for each individual point 
if we look at teh sum of all teh levaerages; then it should be 2 
- its useful ebcause a tpyical leverage value would have be 2/n = mean leverage 
look for hi > 2(2/n) or hi > 3(2/n)
THis is still predictors; 

```{r}
2/28 #Average leverage 
# Dif points, bigger, might have dif influence on model 
2*2/28
3 * 2/28 # These values aren't super sepciacl; they're jsut cut off points for hugher poteint thatl have influence 
#Hat values does this calcutlation for you 

hatvalues(mod1)
```

The data above is ordered; so you can see how the leverages are going to be similar; the ones that are differente will haev a higher leverage, but once it gets cloer, the leverave will be smaller 

Can sort teh values if you want a better idea of what things look like 
```{r}
sort(hatvalues(mod1))

# IFyouw want to look at big leverages 

sort(hatvalues(mod1), decreasing = TRUE)

# Even our smallest and biggest values are not that far away from teh influence values; which are the 2*2/28; etc. 
```
HOw to only pull some of the residual values 
```{r}
head(sort(hatvalues(mod1)), decreasing = TRUE, 10)
```

The above just pulls the biggest 10 so you dont get like 20 pages of residuals on the homework completion 

These values dont appear to have much leverage; but waht about the palm beach example? 
We're going to build up a model so you can redict what people would get in the thing 
```{r}
data("PalmBeach")
ElectionModel = lm(Buchanan~Bush, PalmBeach)
plot(Buchanan~Bush, data = PalmBeach)
abline(ElectionModel)

points(PalmBeach$Bush[50],
       PalmBeach$Buchanan[50],
       col = "red", 
       pch=16) # THe pch function fills in the point; if you want ti more or less filledin ; maybe it changes the visual too,but not sure. 
```

```{r}
2/67 # There are 67 countries 

2*2/67 # Low influence 
3*2/67 # High influnce 

hatvalues(ElectionModel)

head(
  sort(
    hatvalues(ElectionModel),
    decreasing = TRUE, 
    10)
  ) # THis should show 10 results, but only hows a couple ; nto sure what's up 
```

```{r}

PalmBeach$Leverage <- hatvalues(ElectionModel)

PalmBeach[c(13,52,6,29,50,16,48,5,36,53),] # Tehse are the rows that i want to look at; and also teh columsn that i want to look at; no specific columns 

```

Above, want to look at how much the dade country is having on teh thing vs palm beach; it has high leverage so it could also be impacting the model 

```{r}
plot(Buchanan~Bush, PalmBeach)
abline(ElectionModel, col="Blue")

noPalm <- subset(PalmBeach, County!='PALM BEACH')
ElectionModel.noPalm <- lm(Buchanan~Bush, data = noPalm)
abline(ElectionModel.noPalm, col = "red")

noDade <- subset(PalmBeach, County!='DADE')
ElectionModel.noDade <- lm(Buchanan~Bush, noDade)
abline(ElectionModel.noDade, col= "green")

```

**Cook's Distance**
How much would the fit change if one data value were omitted? 
Di increases with eithe rpoor fit (Std. res) and high leverage (hi) 
- Compare to otehr Dis, study any case with Di>0.5; worry if Di>1.0
```{r}
cooks.distance(ElectionModel)
head(
  sort(
    cooks.distance(ElectionModel),
    decreasing = TRUE, 
    8
  )
)

```

```{r}
plot(ElectionModel)
# Cook's distance; 0.5 and 1; we can say if it's outside of this, then it influences the model; tells up impact, less so leverage 
# Why do these points have impaoct and do we want them to impact it? We might have some extreme values that we dont care if they predict or not; good to see if it's nice to see if something is impacting our predictions 
```

NOtes on the 2nd Homework 
- WIll predict things; but you need a set subset; you need to look at a certain subset; code will help; r = eval = FALSE; 
- eval = FALSE; this means that if you run it, it will run, but if youknit it it iwll ignore the chunk; so make sure you get rid of that.