---
title: "STOR 455 - Class 17 - R Notebook"
output:
  word_document: default
  html_document:
    df_print: paged
---
What do we do with categorical variables? 
__Example: Pulse Rates__
- Response Variable:     
		- Y =Active pulse
- Predictors:  
		- X1 = Resting pulse
 	  - X2 = Sex (0=M, 1=F)
- Datat that looks at heart rate and how we can use sex as a categorical indicator varianble and how we should consider it if we add it to our model 
- If we just use sex
--If we wanted to see the difference bt the binary groups, then we could do a two sample t test 

__Categorical Predictor__
Example:  
   Response = Y = Active pulse
   Predictor = X = Sex   
- *Are active pulse rates related to sex? “Usual” procedure?*
- Two-sample t-test (difference in means)
-- How different are the two means vs what I would expect to see from them? 
-- A few ways you can do difference in means test
-- How unlikely would it be that they are this different by chance in our smaples 

```{r message=FALSE, warning=FALSE}
library(readr)

Pulse <- read_csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/Pulse.csv")
head(Pulse)
```

__(using pooled variances), Two-sample T-test for Means__
Ho: mu1 = mu2
Ha: mu1 != mu2

where, (pooled standard deviation):
𝑠_𝑝=√(((𝑛_1−1) 𝑠_1^2+〖(𝑛_2−1)𝑠〗_2^2)/(𝑛_1+𝑛_2−2))

𝑡=(𝑦̅_1−𝑦̅_2)/(𝑠_𝑝 √(1/𝑛_1 +1/𝑛_2 ))
Compare to t with 𝑛_1+𝑛_2−2 d.f.

__R - Two-sample T-test__
- pvalue of 0.004853 is significant

- We want to see if active heart rates show any difference 
- Will add var.equal=TRUE
- By default the dif in mean test assumes that the means are euqla bt the groups and wants to see how unlikely you get this result by chance 
- ytou could also make the assumption that they come from teh same popuilation, so the means and spread should be equal as well 
- We want withthe assumption that the mean active heart rates are equal by sex; 

- Ho: Mean active heart rates = by sex 
- Ha: No equal 

- Evidence there is some difference by sex 
```{r}
t.test(Active~Sex, var.equal=TRUE, data=Pulse)
```

__“Dummy” Predictors__
- We can code a categorical predictor as (0,1) 
- How should this be interpreted in a regression?
- Example:  Y = Active pulse, where 0 = male and 1 = female 

__For summary of modelP__ 
Table, estimate, std, error, tvalue, P 
Intercept, mean for males, 1.330, 63.607, pvalue 
sex, difference for females, 1.874, 2.833, 0.00486
residual standard error....

__t.test(active~Sex, var.equal=TRUE, data = Pulse)__
data:Active by Sex 
t = -2.8329, df = 373, pvalue = 0.004863 <- this is the t-test for significant difference 

sample estimates: 
mean in group 0 mean in group 1 
84.60753 (mean for males), 89.91534 (mean for female)

```{r}
modelP=lm(Active~Sex, data=Pulse) #Active hr predicted by sex 
# In teh data,t eh intercept = mean for the meales ebc its where sex = 0 
# The mean  for 0 = 84 = the male active heart rate 
# The mean for 1 = 89 = female active heart rate, if 
# If out linear model is just an intercept + sex*slope, and sex is a 1 or a 0, then it ends up being intercept + 0 or intercept + slope 
# Getting together = female active heart rate 
summary(modelP)

# looking at t tests
# same pvalue, dif sig digets 
# Can use a linear model like we did d fiference in means test in teh past 
```

__Quantitative + Indicator Predictors__
Example: Y = Active pulse rate
      		X1 = Resting pulse rate
      		X2 = Sex (0,1)
(𝐴𝑐𝑡𝑖𝑣𝑒)̂=8.016+1.165𝑅𝑒𝑠𝑡+2.326𝑆𝑒𝑥

*How do we interpret the coefficient of sex?* 
Ho: B2 = 0
Ha: B2 != 0

With pvalue of 0.116 (because B2 is Sex), There is not evidence to reject the null hypothesis 
and suggest that B2 != 0

```{r}
modelP2=lm(Active~Rest+Sex, data=Pulse) # model with quanti and categorical variables
summary(modelP2)
# Ho: Rest = 0 
# Ha: Rest != 0 
# Small pvalue = reject Ho

# Ho: Sex = 0 
# Ha: Sex != 0 
# Higher pvalue = fail to reject Ho 
```

**Same slope, different intercepts** 
```{r}
plot(Active~Rest, col="blue", data=subset(Pulse,Sex==0))

points(Active~Rest, col="green", data=subset(Pulse,Sex==1))
# plots the points on the graph 

# Below shows where we got the things from 
B_Int = summary(modelP2)$coef[1,1]
B_Rest = summary(modelP2)$coef[2,1]
B_Sex = summary(modelP2)$coef[3,1]

# Plots the line of males and females separately 
lines(
  B_Int + B_Rest * Rest ~ Rest, 
  col = "blue", 
  data = Pulse
  )

lines(
  (B_Int + B_Sex) + B_Rest * Rest ~ Rest, 
  col = "green", 
  data = Pulse
  )
# THis shows that it is forcing us to assume that there is the same slope per sex 
```

__Fit Models to Subsets__
(𝐴𝑐𝑡𝑖𝑣𝑒)̂=9.440+1.1432 𝑅𝑒𝑠𝑡  (Males)
```{r}
Males=subset(Pulse, Sex==0)
modelPM=lm(Active~Rest, data=Males)
summary(modelPM)
# Intercept 9.43 and a slope of 1.1432 for the male model 
```

__Fit Models to Subsets__
(𝐴𝑐𝑡𝑖𝑣𝑒)̂=9.153+1.1823 𝑅𝑒𝑠𝑡  (Females)
```{r}
Females=subset(Pulse, Sex==1)
modelPF=lm(Active~Rest, data=Females)
summary(modelPF)
# Intercept of 9.1527 witha  slope of 1.1823 for the female model 
```
We see some difference between the output.  Is this a signifigant difference or would I just expect to see this by chance? 

__Plotting the lines__
- Are these lines “significantly” different?
```{r}
plot(Active~Rest, col="blue", data=subset(Pulse,Sex==0))

points(Active~Rest, col="green", data=subset(Pulse,Sex==1))
# The above code puts the dots on the graph 

# the below code puts the line of the models with male and female on the graph

lines(
  summary(modelPM)$coef[1,1] + summary(modelPM)$coef[2,1] * Rest ~ Rest, 
  col = "blue", 
  data = Pulse
  )

lines(
  summary(modelPF)$coef[1,1] + summary(modelPF)$coef[2,1] * Rest ~ Rest, 
  col = "green", 
  data=Pulse
  )

# Now we are working with 2 different models instead of one 

# The slopes are slightly different, and we are allowing for the different rate of change; we want to do this with the entire dataset. 
```
Y = B0 + B1X1 + B2X2 + B3X1*X2+Error
- We want to create a line that by changing the value of one indivator variable, we can change what teh intercept of the prediction is, as well as what the slope of that model is 
- The interaction term does this 
- When sex = 1 what should slope be and when sex = 0 what should slope be? 
- When sex = 0, then the some things go away, but when sex = 1, the B2 Term will effect the intercept of the model and B3 will affect the slope of the model 

__Comparing Two Regression Lines (with a multiple regression)__
- Example:   
	- Y=Active pulse
	- X1= Resting pulse	
	- X2= Sex(0,1)

𝑌=𝛽_𝑜+𝛽_1 𝑋_1+𝛽_2 𝑋_2+𝛽_3 𝑋_1 𝑋_2+𝜀
Y = Intercept + Quantitative + Indicator + Interaction

__Quantitative + Indicator +Interaction__
(𝐴𝑐𝑡𝑖𝑣𝑒)̂=9.440+1.1432𝑅𝑒𝑠𝑡−0.287𝑆𝑒𝑥+0.039𝑅𝑒𝑠𝑡∗𝑆𝑒𝑥
- How does this relate to the two lines? 
- Substitute Sex=0 and Sex=1

```{r}
# Interaction terms
# ANOVA Assumptions 
# Ho: All Bi = 0 
# Ha: At least one Bi != 0 
modelPint=lm(Active~Rest+Sex+Rest*Sex, data=Pulse)
# BEcause rest is sig, it doesn't appear that the intercept change is useful for us to do 
# interaction bt rest and sex has a high pvalue, tells us we might not need the interaction term 
summary(modelPint)
```

*How can we make a sig test for this?* 
- When we dont appear to have evidence for sig difference in slope or intercept? 
- Are there different lines to predict what we want? 
- If there are, then the last two coeff would be 0 

__Tests to Compare Two Regression Lines__
Y = Bo + B1X1 + B2X2 + B3X1X2 + Error

- **Different Slope** - T test
-- Ho: B3 = 0 
-- Ha: B3 != 0 

- **Different Intercept** - T test
-- Ho: B2 = 0 
-- Ha: B2 != 0  

- **Different lines** - See Multiple Regression Model section below
-- Ho: B2=B3=0
-- Ha: B2 != 0 or B3 != 0 

__Multiple regression model__
- **Testing one term at a time:**
-- T-test
-- Ho: B1 = 0 
-- Ha: B1 != 0 

- **Testing all terms at once**
-- ANOVA
-- Ho: B2=B3=0
-- Ha: Some Bi != 0 

*Is there anything in between?* 

__Nested Models__
- **Definition:** If all of the predictors in Model A are also in a bigger Model B, we say that Model A is nested in Model B.

- Example:  𝐴𝑐𝑡𝑖𝑣𝑒=𝛽_0+𝛽_1 𝑅𝑒𝑠𝑡+ 𝜀   is nested in 
            𝐴𝑐𝑡𝑖𝑣𝑒=𝛽_0+𝛽_1 𝑅𝑒𝑠𝑡+_2 𝑆𝑒𝑥+_3 𝑅𝑒𝑠𝑡∗𝑆𝑒𝑥+𝜀 

- *Test for Nested Models:* 
- Do we really need the extra terms in Model B?
- i.e. How much do they “add” to Model A? 
```{r}
# Something in between 
# ANOVA = all coef are zero vs at least one is nonzero - we compare to null mode, how much do we explain vs a null model? 
# We can sub a different model to the null model 
# Caveat: The sub model has to have a nested subset of what we are working with in the bigger model 

modelP_Reduced = lm(Active~Rest, data=Pulse)

# NEsted F Test 
# Do we need all these things in the model? 
# IF we add things tothe model we may get a smaller mallow CP, but is it signifigant improvement? This tells you 
anova(modelP_Reduced, modelPint) # Nested test 
# First line = test compare model to null model, and its sig 
# THen build a model with rest and sex in it and comparing the model before to this model 
# To see if addingteh sex predictor increases teh varaibility signifigant;y (It is not in this explame)
# Third line compares rest, sex, and interaction to just the model with rest and sex in it
# Tells you if we are explaining an extra amount of the varibility by adding the interactiont erm (This tells you you are not explaining a good amount extra) 
# Only showed up this way ebcause fo the order you put it in ANOVA, if you change the order, you change the order it analyzes things and it might change what it says 

# F test stat is calc similiarlly, the dif is when we think about sum of squares; its not how much teh SS in this model it's how much they are in this model taking away what is in the reduced model 
# How much variability is being explained by these extra things, not the whole model as a whole 
```

```{r}
anova(modelPint)
```

__Nested F-test__
- Basic idea: 
1. Find how much “extra” variability is explained when the “new” terms being tested are added. 
2. Divide by the number of new terms to get a mean square for the new part of the model. 
3. Divide this mean square by the MSE for the “full” model to get an F-statistic. 
4. Compare to an F-distribution to find a p-value.

- Test:   Ho: Bi = 0 for a “subset” of predictors
          Ha: Bi != 0  for some predictors in the subset

- 𝐹=(((𝑆𝑆𝑀𝑜𝑑𝑒𝑙_𝐹𝑢𝑙𝑙−𝑆𝑆𝑀𝑜𝑑𝑒𝑙_𝑅𝑒𝑑𝑢𝑐𝑒𝑑))⁄(# 𝑝𝑟𝑒𝑑𝑖𝑐𝑡𝑜𝑟𝑠))/𝑀𝑆𝐸𝐹𝑢𝑙𝑙
- F = ((Explained by full model - explained by reduced model)/#predictors tested in Ho)/MSEFullthat is based on the full model 
- Compare to F-distribution 

__Nested F-test__
- 𝐴𝑐𝑡𝑖𝑣𝑒 =𝛽_0+𝛽1𝑅𝑒𝑠𝑡+_2 𝑆𝑒𝑥+ 3𝑅𝑒𝑠𝑡𝑆𝑒𝑥 +𝜀 
- H0: β2=β3=0
- Ha: Some βi != 0 
- Compare mean square for the “extra” variability to the mean square error for the full model. 

```{r}
modelPint2 = lm(Active~Sex+Rest+Sex*Rest, data = Pulse)
anova(modelPint2)
# First one being done is sex a sig pred of active heart rate? 
# Yes it is, but it looks different than the other tests because they were doing soemthing different 
# This test is after considering teh variability explained in teh active heart rat eby a resting heart ratel is sex sig after that? 
# We have a big p value so its not, 
#If we dont take into account the varability in account by resting heart rate first, then sex alone is useful 
# ADding rest and then sex adn rest together has a pvalue of 2.2 to the -16, to its a sig amount explained 
# The last row is teh same because we are still comparing teh same test as we are before 
# A mdoel wtih all 3 vs a model withjust sex and rest 
```
```{r}
modelPint3 = lm(Active~Rest*Sex, data = Pulse)
summary(modelPint3)
# Even though its jsut an interaction term, R assumed we wanted each indidivual terms as well 
# YOu dont want an interaction without each of the terms in teh model before that 
# R knew that I should do that, so it gave it to you 
```
ANOVA Tests 
- We dont want ot just put one function into it becase thatis depend on the order of the predictors of the model 
- ONly lets us test one predictor at a time 
- Adds one layer at a time 

We want to compare a model with all 3 predictors with just 1 predictor (just restin gheart rate) 
```{r}
modelP_Reduced = lm(Active~Rest, data = Pulse)
summary(modelP_Reduced)
```

- Then we wnat to do the ANOVA on the reduced and others 
- To see how much variability is being explaine dby adding those new terms to it 
```{r}
anova(modelP_Reduced, modelPint)
# This tells us the df is 2, whih tells us tehre are 2 variables different between tehse two models 
# We get an idea of teh variability explained bt these two models and see tha tthe SS (the 512.14) = the amount of varabiltiy explained by adding teh sex and interaction term to the model ; that is a small amount in the long run and that is why the f test stat is small and the p value is high 
# This lets you test both those terms together 

# BEnefit: We are jsut comparing two extra terms; if there were 10 terms different, then we could just do tests of each 10 terms individually, but tehn we'll run into more error issues 
# When we do it all at once, we get a big pitcure if there are any differences and if we need to we can investigate further to see wehre teh differences are 
```
Nested tests tell you if you're looking at noise or if you're looking at something signifignat 

