---
title: "STOR 455 - Class 20 - R Notebook"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r message=FALSE, warning=FALSE}
library(readr)
library(leaps)

StateSAT <- read_csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/StateSAT.csv")

source("https://raw.githubusercontent.com/JA-McLean/STOR455/master/scripts/ShowSubsets.R")
```
__Example:  State SAT__
- Model #1:  Y=SAT   vs. X=Takers

```{r}
mod = lm(SAT~Years+Expend+Rank, data=StateSAT)
summary(mod)

plot(mod)
# We have a curve in teh residual plot 
# The noramility is an issue 
# The one state has a lot of influence, WE think it's AK, that has a fewer precentage of the population in public schools 
```

```{r}
plot(SAT~., data=StateSAT[2:8])
# Here, we see the correlation between each of the variables in the dataset (each of teh numerical variables in teh dataset) 
# Expend is iffy 
# Years is good
# Takers looks like the best pattern, but we dont have it because it doesn't look lienar 
```

```{r}
# LEts look at just the taker's variable 
modSAT1 = lm(SAT~Takers, data=StateSAT)
summary(modSAT1)
plot(modSAT1)
# Small pvlaue and high variability is described 
# Linearity condition is super messed up though, so what can we do?  - Transofmrations! 
```

__Polynomial Regression__
For a single predictor X:
ð‘Œ=ð›½_ð‘œ+ð›½_1 ð‘‹+ð›½_2 ð‘‹^2+â‹¯+ð›½_ð‘ ð‘‹^ð‘+ðœ€

- **LINEAR** 
-- ð‘Œ=ð›½_ð‘œ+ð›½_1 ð‘‹+ðœ€

- **QUADRATIC** 
-- ð‘Œ=ð›½_ð‘œ+ð›½_1 ð‘‹+ð›½_2 ð‘‹^2+ðœ€

- **CUBIC** 
--ð‘Œ=ð›½_ð‘œ+ð›½_1 ð‘‹+ð›½_2 ð‘‹^2+ð›½_3 ð‘‹^3+ðœ€

__Issues with Polynomial Regressionn__
- We can move it up and down based on the intercept or make it widder or thinner based on the slope;
- We can't change where the vertex is

```{r}
#What if we raise taker's the the 2nd power? 
# We have to insolute it so R will actually do it 
# THis made it worse, so to shift it to the right, we have to use a quadratic regression line  
modSAT2 = lm(SAT~I(Takers^2), data=StateSAT)
summary(modSAT2)
plot(modSAT2)
```

__Polynomial Regression in R__
- We can add as many powers we want but then we might be over fitting, so that's not always best

Method #1: Create new variables with predictor powers.
- Create a new model of takers^2
- Use to shift to the right or left
```{r}
StateSAT$TakersSq = StateSAT$Takers^2
# ameks a new column of takers^2

modSATquad1 = lm(SAT~Takers + TakersSq, data=StateSAT)
summary(modSATquad1)
# Above, makes a model with takers^2
# We see that it's a pretty sig model 
# It looks good, but does it help with the residuals? 

plot(modSATquad1)
# The linearity looks pretty good 
# Constance variance could be better because we dont have a lot of data 
# the normal, looks pretty good too 
```

__Polynomial Regression in R__
Method #2:  Use I( )in the lm( )
- Does the same thing as a bove, but it does it with just the insulate function
```{r}
# Quadratic model for SAT
# (ð‘†ð´ð‘‡)Ì‚=1053.1âˆ’7.1616ð‘‡ð‘Žð‘˜ð‘’ð‘Ÿð‘ +0.0710ã€–ð‘‡ð‘Žð‘˜ð‘’ð‘Ÿð‘ ã€—^2
modSATquad2 = lm(SAT~ Takers+ I(Takers^2), data=StateSAT)
summary(modSATquad2)
```

__Polynomial Regression in R__
Method #3:  Use poly
- Does the same thing as the other methods, but it just tells it to make a polynomial
- This will be treated as one unit instead of separately 
```{r}
modSATquad3 = lm(SAT~poly(Takers, degree=2, raw=TRUE), data=StateSAT) # 2 = quadratic 
summary(modSATquad3)
# Same values 
```

```{r}
#ANOVA TREATS THE DIFFERENT METHODS DIFFERENTLY
anova(modSATquad1)
# Looks at takers vs takers^2
# tells you adding the squared term is useful for us 
anova(modSATquad3)
# Condesnes the terms into one 
# Jsut a test on the one model

# Doing the same things, but the function treats it differently depending on the method you use 
```

```{r}
# Quadratic model for SAT
plot(SAT~Takers, data=StateSAT) # Plot raw data

# Pull out teh coeff for the terms for the quadratic model 
B0_modSATquad2 = summary(modSATquad2)$coef[1,1]
B1_modSATquad2 = summary(modSATquad2)$coef[2,1]
B2_modSATquad2 = summary(modSATquad2)$coef[3,1]

# curve(INtercept, coef*x, coef*x^2, add = TRUE)
curve(B0_modSATquad2 + B1_modSATquad2*x + B2_modSATquad2*x^2, add=TRUE)
# Looks like it fits really well 
```
Would a Cubic work better? 
```{r}
#Cubic MOdel
modSATcubic = lm(SAT~ Takers+ I(Takers^2) + I(Takers^3), data=StateSAT)
summary(modSATcubic)
# R and rsqaured are simular 
# Models of dif predictors, adj r squared is better measure 
# Its a little worse than a 2 model 
# High p value 
# tells us that not as sig 
```

```{r}
# Cubic MOdel
plot(SAT~Takers, data=StateSAT, main="Cubic Model")

B0_modSATcubic = summary(modSATcubic)$coef[1,1]
B1_modSATcubic = summary(modSATcubic)$coef[2,1]
B2_modSATcubic = summary(modSATcubic)$coef[3,1]
B3_modSATcubic = summary(modSATcubic)$coef[4,1]

curve(B0_modSATcubic + B1_modSATcubic*x + B2_modSATcubic*x^2 + B3_modSATcubic*x^3, add=TRUE)
# Doesn't look super differnet 
# In the end, there's not a lot of change witht eh cube term 
# IT has a small coef compared to the others as well, so so not super big influence out the gate but we dont know if its a tually influencital we would haev to check other htings
```

```{r}
anova(modSATcubic)
# Tells you if all the other models are sig
# Tells us that takers to the 3rd isn't useful 
```

```{r}
car::vif(modSATcubic)
# shows that there is a high multicollinearity with takers
```


```{r}
modSATquad4 = lm(SAT~ Rank+ I(Rank^2), data=StateSAT)
summary(modSATquad4)
# Model looks pretty good, rank and rank^@ are not sig; btut ehre is a hug amount of multicollinearity 
# P value is close to 0, but the individual tests say things different because of the multicollinearity 

plot(modSATquad4)
# Residual anaysis isn't too bad
# Normal is a little bit of an issue
 
plot(SAT~Rank, data=StateSAT)
# Slight curve when we raise to a power 
# Not a linear relationship 
# We could fit a line to it, but there may be some issues 
# THis is the more ideal situation than a line 

B0_modSATquad4 = summary(modSATquad4)$coef[1,1]
B1_modSATquad4 = summary(modSATquad4)$coef[2,1]
B2_modSATquad4 = summary(modSATquad4)$coef[3,1]

curve(B0_modSATquad4 + B1_modSATquad4*x + B2_modSATquad4*x^2, add=TRUE)
```

```{r}
# IF we jsut used a squared rank term, then we would 
# IT ssimilar, but not quite just right
# WE have to flatten the parabola out more 
# Still centered at the zero, jsut stretched further and its less useful 
plot(SAT~Rank, data=StateSAT)
mod2 = lm(SAT~I(Rank^2), data=StateSAT)

B0_mod2 = summary(mod2)$coef[1,1]
B1_mod2 = summary(mod2)$coef[2,1]

curve(B0_mod2 + B1_mod2*x^2, add=TRUE)
```

__Polynomial with one predictor__ 
- We can use different order models that look at other models in a 3D space for predictors 

__Second Order Models__
Definition: A second order model for two quantitative predictors would be
ð‘Œ=ð›½_ð‘œ+ð›½_1 ð‘‹_1+ð›½_2 ð‘‹_2+ð›½_3 ð‘‹_1^2+ð›½_4 ð‘‹_2^2+ð›½_5 ð‘‹_1 ð‘‹_2+ðœ€
Y = INtercept + First Order + First Order + Quadratic + Quadratic + INteraction, where Wuadratic+ INTeraction = Second order 

Example: Try a full second order model for 
Y=SAT using X1=Takers and X2=Expend


