---
title: "STOR 455 Class 3 R Notebook"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---

```{r message=FALSE, warning=FALSE}
# message=FALSE, warning=FALSE supress warnings and messages from appearing in knitted html

library(readr)

DistanceHome <- read_csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/DistanceHome.csv")

# If notebook and csv file are saved in the same folder
# DistanceHome <- read_csv("DistanceHome.csv")

Domestic=subset(DistanceHome,Distance<250)
```
__Single Quantitative Predictor Model__ 
- Notation:   
-- Y = Response variable
-- X = Predictor variable

*Assume (for now) that both Y and X are quantitative variables.* 
Y = f(x) + Error 

_Simple Linear Model_
- X = Single quantitative predictor
- Y = Quantitative response

*Goal*: Find a line that best summarizes the trend in the data.

Y = Bo + B1x + Error 
Response = Intercept + Slope*Predictor + Random Error 
*Assumptions:* 
- Assume:  Error ~ Follows a normal distribution and independent
- There are 3 parameters to estimate: Bo, B1, and std error

__Scatterplot in R__ 
*See below* 
```{r}
plot(Distance~Hours, data=Domestic)
```

__Least Square Regression in R__
*Syntax:* Syntax:    lm(Response~Predictor,data= )

```{r}
lm(Distance~Hours, data=Domestic)
```

_R – Plot with Fitted Line _
```{r}
mod1=lm(Distance~Hours, data=Domestic)
plot(Distance~Hours, data=Domestic)
abline(mod1)
```


_Simple Linear Model- Conditions_ 
**Model:**
1. Linearity: The means for Y vary as a linear function of X.
**Error:**
2.Zero Mean: The distribution of the errors is centered at zero.
3.Constant variance: The variance for Y is the same at each X. (Homoscedasticity)
4.Independence: No relationships among errors.
5.Normality:
- Residuals are normally distributed
- (sometimes) At each X, the Y’s follow a normal distribution.

_Linear_ 
Look for consistent curvature or non-linear patterns

_Constant Variance_ 
Look for “fan-shaped” pattern
- Fan-shaped is **bad** 

```{r}
summary(mod1)

mod1$residuals
mod1$fitted.values

plot(mod1$residuals~mod1$fitted.values)
abline(0,0)
```

_Residuals_ 
Look at a histogram of the residuals 
Look for clear skewness and outliers
- skew and outliers are **bad**

```{r}
hist(mod1$residuals, breaks=10)
```

_How much Variability is Expected?_

```{r}
qqnorm(mod1$residuals)
qqline(mod1$residuals)
```

```{r}
#Sets randomization
set.seed(455001)

#Sample of 99 values from a Normal distribution; mean=0; sd=1; sorted ascending
x = sort(rnorm(99,0,1))
hist(x)

#list of integers 1 through 99.
y = c(1:99)

#z-scores of dataset of 99 values if perfectly normally distributed
normy = qnorm(y/100)

plot(x~normy)
abline(0,1)
```

```{r}
x <- rnorm(54, 0, 18.26)
qqnorm(x)
qqline(x)
```
