---
title: "STOR 455 Class 12 R Notebook"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r warning=FALSE, message=FALSE}
library(readr)
library(Stat2Data)
library(car)

data("Houses")

StateSAT <- read_csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/StateSAT.csv")

source("https://raw.githubusercontent.com/JA-McLean/STOR455/master/scripts/anova455.R")
``` 

```{r}
head(Houses)
cor(Houses)
```

```{r}
HouseModel=lm(Price~Size+Lot,data=Houses)
# Linear model that predicts jprice by size and lot of the house.

summary(HouseModel)
# Tests teh coef size and lot are equal to zero; that they do not have a realtionship with price 
# Alternative: that at least one of those is nonzero 
# Very low pvalue, very unlikly that we would get this sample if the nuill was true; we have evide nce to say that at least one of these has a non zero slope 

# Whtat is the FTest stats and what it is useful? 
# The pvcaalue is really what we want
# The f tests stat, if its big or small depends on the sample size and the number of predictors; 
# GFOr a small sample of small predictors, 10 = big number 
# Large sample wiht lots of predictors, 10 = small number 
# Mostly focus on the pvalue and how to interpret that 

# Looking at the coeff table 
# The Ho: Coeff of size = 0
# Ha: Coef size != 0 
 # Same thing for lot 
# Each ahas an indivual test sthat ehre is probably not evidence of a relationshipo 
# The two appear to be contradictory there 
# When you're writng out the hypotehsis, you can say in words what he ahsa said or you can write it our mathmatically
```

```{r}
#cor.test(Houses)

cor.test(Houses$Size, Houses$Price)
# jfThe relationship bt price and size are with a low pvalue, 0.0008; havbe evidence of a relationship here that is non zero correlation 
# the same test with the other; there is no evidence of ra realtionship 
cor.test(Houses$Lot, Houses$Price)
cor.test(Houses$Size,Houses$Lot)

# We get contradcitory results because of multicollinearity 
# there is an issue wher ethe variance is being inflated when we do these tests 
# IF we loook at the relationship between teh predictors (Lot and size) we see that eh correaltion is really high 0.7686 ish (Not exaclty) it's a signifigant realtionship 
# This is driving the conficting results because too much is being explained by the two thing s
# It's not inherantly bad, it's not telling ust htat there is no realtionshipship, it's just saying that we dont haev evidence to say ther eis s asignfigiant realtionship 
# If we have a lto of predictors that are highly correlationed you might not want to use them all on our model 
# If these predcitors are explaining the same thing, then why include both? IT sjust going to cause problems 
# THis can cause overfitting problems 
```

*Simple models are idea, than overaly complicated ones* 

__Multicollinearity__

- What is it? 
-- When two or more predictors are strongly associated with each other.
- Why is it a problem? 
--Individual coefficients and t-tests can be deceptive and unreliable. 

*NOPtes* 
- Makes the tests deceptive and we need to know that there is multicoloinarity going on 
- More its unrealiable tests if there are multicollinearity 
- It makes it harder to interpret but it means that our model acna be simpler than what we have 
- so it really means, jsut change you rmodel a little 

__Effects of Multicollinearity__
- If predictors are highly correlated among themselves:
1. The regression coefficients and tests can be extremely variable and difficult to interpret individually.
2. One variable alone might work as well as many.

```{r}
anova(HouseModel)
```

```{r}
anova455(HouseModel)
```

*NOtes* 
-0 WE can see where the correlation is bt lot and size 
- One way we can test is to see how closely correlated things are
-- THis is fine whwen you're jsut looking at two things; if we have a lot of predictors then teh correlation between things cna be hard to use as a measure ebcause there are more things to look at 
- Solution: Build a new model for each predcitor where th remaining predcitors are the predictors of that model
- In this case, we could build a model for the size of a house and use the rest of the predcitors as predictoyrs (Would just be lot in this case) and do teh same hting for lot and make a model where size is the predictor for that 

*See below* 
```{r}
mod=lm(Size~Lot, data=Houses)
summary(mod)
# We want ot look at the multiple r sqaured; it says that almost 59% is beign predictoed by how big the lot is 
# Thats just the correlation squared,. the .77 squared; when we get the bigger models, we are going to have to do more to calcualte that 
# We see how much variability is explained there by the two predicotrs 

# can use this to se ehow much the variance is being inflated 
# The variance that is being calcualted that is for each paredictor is not done in isolation; its taking into account the other predictors; more multicollinearity will increase the variance 
summary(mod)$r.squared
# how to pull out the multiple r-squiared from the model 
```

__How do we detect multicollinearity?__
1.  Look at a correlation matrix of the predictors.

```{r}
round(cor(Houses), 2)
```
2.  Compute the Variance Inflation Factor (VIF). 
- (Beware if VIF > 5)
- where Ri2 is for predicting Xi with the other predictors.
- 𝑉𝐼𝐹 > 5 or 𝑅𝑖2 >80%

```{r}
# How to account for the inflated variance in places with possible multicollinearity 
VIF = 1/(1-summary(mod)$r.squared)
VIF
# If VIF is 5 or more, then ther emight be a lot of multicollinearity going on
# This would mean the adjusted r sqaured would be above 80 or more 
# We are saying that the variance is being aadjusted by a factor of 2.42
# We get the 2.42 by the VIF 

# If we look at the summary of the housemod 
# the variance of size and the stderror = 17.7, when we are doing a hypothesis test for the slope of size, then we are caclauting a t stest stat - the estimate for slope/Stderror; 
 # 23.2/17.9 = 1.313 which is the tvalue 
# that's where that tvalue is coming from 
# We could pull it our better with a summary funciton, but we're not going to 
# So this outcome is about 1.31 stdar devations away if we didnt have a relation between teh things if there was no realtion 
# The variance that we used in this calvcualtion, because of the multicollinearity is being inflated by this facotr 
# This is teh variance inflation factor and we are calculting the stadard error 
# Std = sqrt(variance) 

#Go down to sqrtt(VIF) code
```

__Finding VIF with R__
1. 1. Brute force. 
     Fit a model to predict Xi using the other predictors and find 𝑅𝑖2.
- Compute: 𝑉𝐼𝐹=1/(1−𝑅𝑖2)
- Example: Find VIF for Size when using Lot to predict Size

```{r}
summary(HouseModel)

sqrt(VIF) # This is how much that variance is being inflated 
# Look at the summary of the houses model 
summary(HouseModel)$coeff[2,2]/sqrt(VIF)
# sqrt(VIF) = how much its being inflated 
# If no correlation., then we would haev a variance of 11.1; so now if were to divide the slope by this value instead 

summary(HouseModel)$coeff[2,1]/summary(HouseModel)$coeff[2,2]/sqrt(VIF) # This is what we would get for the slope if we didn't have the multicollinearity 
# So multicollienarity has a really big impact on the model 

# We dont have to do the math above in practice, it's really just to learn how and why the infaltion is affecting teh table 
```

__Finding VIF with R__ 
2. 2. Install car package
- use  vif( )function or use VIF.R script from Sakai
*See below for installing car package and using the vif function* 
```{r}
vif(HouseModel)
# This is how to do what you did above, but really short form 
# This is what you would use to look at the inflation 
# This will be the same when you are lookign at t athing with 2 predictors 
# Ity will be different when you ahev mutliple predictors 

# Even though it changes our result in the data from a sig to a non sig realtionship; multilcolinearity wise, it snot a huge realtionshio
# mostly because its  samll dataset 
```

__What to do if you’ve got Multicollinearity?__
1.Choose a better set of predictors
2.Eliminate some of the redundant predictors to   leave a more independent set.
3.Combine predictors into a scale.
4.“Ignore” the individual coefficients and tests.

Note: Predictions aren’t necessarily worse if some predictors are related – it’s just conclusions about individual terms that might be confused.

**NOTES** 
- Looking at a bigger dataset when it's not so straightforward to know if there is multicollinearity or not

__Example: State SAT Scores__
Source:    Statistical Sleuth, Case 12.1 pg. 339  
Response Variable:     
      SAT    =Average combined SAT Score
Potential Predictors:  
     Takers  = % taking the exam
     Income = median family income ($100’s)
     Years    = avg. years of study (SS, NS, HU)
     Public   = % public school
     Expend = spend per student ($100’s)
     Rank     = median class rank of takers

```{r}
head(StateSAT)
# Some states might have lower SAT takers ebcause the ACT mgiht be better 
```

__Example: Predicting State SAT__
Data:   StateSAT   
Response:  SAT
Possible Predictors: 
	Takers	Income	Years	Public	Expend	Rank

Find the “best” model for GPA using some or all of these predictors. 

*What determines “best”?*

```{r}
plot(SAT~., StateSAT[,2:8])
# Want ot predict the average SAT per state 
# The ~ will take all the remaining columns 

```

```{r}
SAT_Model = lm(SAT~., data=StateSAT[,2:8])
plot(SAT_Model)
summary(SAT_Model)

# We see some weird things are happening where there ar ea lot of NAs happenign here 
#It's because we have categorical variables and r is trying to get those results in number form of us 
# We want to use only the nnumeric things 
# That is what this model does above 

# The bottom line does a n anova test witht eh assumption that the slope for the columsn are all zero and there is no relation bt sat scores vs the alternative that a tleast one is non zero in thei model 
# We Think base don this model at least one is a good predictor 
# IF we look at the individual tests for slope, we see where 3 of them have low pvalues (YEars, Expend, and Rank) Where the p value is low; this model those seem to have a strong relationship with SAT score 
# These results might be decieving if there are multicollinearity 
# there is most likely multicollinearity here because its all interfomration fro one state 
# There is probably some lurking variables there that are in the background making these things highly correlated to each other 
```

```{r}
vif(SAT_Model)
# looking at the VIF for this model, we see that Takers and Rank have really high inflation rates, which means they should probably be expcluded from the final regression model 
# It teslls us wif we make a mdoel with takers as the repsonse, we will get a really high r squared value; almost all the other varibaility in takers is being expained by the other varianceles 
# we probably dont need takers in teh model if everythign else is already doing that for us 
# We could probaly not need rank either as well for the same reason 
# Or maybe just rank or just takers is all we need to predcit.
# It gives us some informatoin, but the biggest thing is we are skipping a good ifrst step 
# Does it meet the lienar model conditions? 
```

*Does it meet the linear model conditions?* 
- Look at residual analysis of the data 
- Does this data meet the criteria, and if it doesn't, where are these problems occuring? 
- Becaufore we had 1 variable and wanted to do a transfomration, we could jsut ransforma the predictor/response
- Now we have ore predictors and some might have lienar issues while others don't 

```{r}
plot(SAT_Model)
# We can look at these plots to see if the ocnditions are met 
# The model doesn't appear to fit a line well 
# Transfomration might be useful 
# Or, dont include all the variables in teh model 
# Maybe tranfomraiotn or may less predicotrs 
qqPlot(SAT_Model)
# normalitiy is a big issue 

# COnstancce variance: Not a big issue 
# Leverage/Cook's distance - super wonky looking, probably wouldn't trust it right now 
# One data point has high influence in model, should probably look at state 29 because of teh cook's plot
```


```{r}
# WE could also look at which varibales are probkematic for us
plot(SAT~., data = StateSAT[,2:8])
# Tehse plots will help us see wehre things might be a good fit for choosing a predictor 
# Will also help show where we might want to look towards what has outliers 
# first is SAT by takers - there is a curve we could use; does a transformation help this? not so much , but some others might 
# Takers seems a pretty good predicotrorl; would haev to worry about lower states 

# Income, doesn't look like a good predicotr, but it is appear to have a conneciton somewhere; but it's not the best; not super linear and not as clear as other s
# YEars: nothing jumps out, but it's hard to see a pattern, there is something there 
# Public, its hard to say what is going on, that is one that is different than the rest, this is probably the state 29 that has high influence; this variable is probably messing up our data 
# Expend; same issue with one state is apearing to spend more omoeny than the rest of the state; one point has a lot of influence 
# Rank: This is pretty definded realtionship; not a line, but appears to be a good varibale here 
# Guessing: The high VIF bt Takers and RAnk; they appear to expain similar amount of varibility withteh SAT scores 
```
- We have all tehse ariables; how do we make the best model? 
- We could go on teh r-squared alone, then it's pretty good; but you should check the condiotns and that makes it a sus model 
- Different ransfomraitons could make the model better and make the model better 

- need to see the realtionships to see if there are different combos that will give a better linear conditoins and realtionship between teh model predictors 


__Predictor Selection Methods__
- **ALL ARE SET ON THE CONDITIONS BEING FINE, IF CONDITIONS ARE NTO FINE THAT IS BAD** 
Choosing an effective set of predictors
- All subsets (All combinations of predicotrs in the model and compare all to each other; there is a certain point in which you cant really do this on a compauter because its really hard on a compauter)
- Backward elimination 
- Forward selection
- Stepwise regression
