---
title: 'Group Project #3'
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
SleepStudy <- read_csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/SleepStudy.csv")
head(SleepStudy)
```
1) Construct a logistic model to predict the _Stress_ of a student using _GPA_, _AverageSleep_, and _ClassYear_ as the predictor variable.

```{r}
mod1 = glm(factor(Stress) ~ GPA + AverageSleep + ClassYear, data = SleepStudy, family = binomial)
summary(mod1)
```


2) For a sophomore student that has a 3.50 GPA, gets an average of 8 hours of sleep a night, what does your model predict is their stress level? Interpret.

```{r}
'The student has a 0.75 chance of having a high level of stress.'

student = data.frame(GPA = 3.50, AverageSleep = 8, ClassYear = 2)
predict(mod1, student, type="response")
```
3) Construct a second logistic model to predict the _Stress_ of a student using _GPA_, _AverageSleep_, and _ClassYear_ as well as the interaction between _GPA_ and _ClassYear_ as the predictor variable.

```{r}
mod2 = glm(factor(Stress) ~ GPA + AverageSleep + ClassYear + GPA*ClassYear, data = SleepStudy, family = binomial)
summary(mod2)
```


