---
output:
  word_document: default
  html_document: default
---
# Logistic Regression
### Building a Model with a single predictors

```{r}
library(readr)
sleep = read.csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/SleepStudy.csv")
```

1)    Construct a model predicting _AllNighter_ by _ClassesMissed_. Include a summary of this model. Use this model to perform a hypothesis test to determine if there is significant evidence of a relationship between _AllNighter_ and _ClassesMissed_. 
```{r}
allnighter = glm(AllNighter ~ ClassesMissed, family = binomial, data = sleep)
summary(allnighter)
```
Hypothesis test
```{r}
anova(allnighter, test = "Chisq")
```
Null hypothesis: The slope to the relationship between _AllNighter_ and _ClassesMissed_ is zero.

Alternative hypothesis: The slope to the relationship between _AllNighter_ and _ClassesMissed_ is nonzero.

2)    For a student that has missed 9 classes, what is the probability the model predicts that the student has pulled an all nighter?
```{r}
sleepmod = glm(AllNighter~ClassesMissed, data=sleep, family=binomial)
predict(sleepmod, data.frame(ClassesMissed = 9), type="response")
```














