---
title: "Stor455-Hw9"
output:
  word_document: default
  html_document: default
---

```{r}
library(readr)
library(Stat2Data)
```

```{r}
data("Election16")
head(Election16)
```
1. Run a logistic regression model to predict _TrumpWin_ for each state using the per capita _Income_ of the state. Print a summary of the model. 

2. Find a 95% confidence interval for the odds ratio using the model constructed in question 1. (Explain if you wanna)

\newpage





Answers

1. Run a logistic regression model to predict _TrumpWin_ for each state using the per capita _Income_ of the state. Print a summary of the model. 

```{r}
Election_logitmod = glm(TrumpWin~Income, 
                        data = Election16, family = binomial)
summary(Election_logitmod)
```


2. Find a 95% confidence interval for the odds ratio using the model constructed in question 1. (Explain if you wanna)

```{r}
exp(confint.default(Election_logitmod))
```
We are 95% confident that for every increase of one dollar, the person's odds that Trump Wins changes by a factor between 0.999694 and 9.999127e-01. 





