---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---
Group Members:
Katherine Bacon
Kay Youngstrom
Ali Floyd
Bethany Newcomb

```{r}
library(readr)
sleep <- read.csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/SleepStudy.csv")
```
```{r}

```

8. Interaction Terms

Create a model that predicts GPA based on Average Sleep.

```{r}
mod1 = lm(GPA~ AverageSleep, data = sleep)
summary(mod1)
```

Now create a model precting GPA based on Average Sleep, Classes Missed, and an interaction between these two variables.

```{r}
mod2 = lm(GPA~ AverageSleep + ClassesMissed + AverageSleep*ClassesMissed, data = sleep)
summary(mod2)
```

Does the ClassesMissed term improve our model? Use Anova or another test we have learned in class. No need for a formal hypothesis.

```{r}
anova(mod1, mod2)
```


The P-value is smaller than .05 which shows that it is significant. This means that at least one of the coefficients of ClassesMissed is non-zero and therefore makes our model better.









