---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

```{r}
library(readr)

Sleepstudy = read_csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/SleepStudy.csv")
```
```{r}
Sleepstudy
summary(Sleepstudy)
```
1) Construct a two-way ANOVA model for the mean GPA in  the dataset, _GPA_, using the _LarkOwl_ and _Stress_ as the predictors. Include the output showing the ANOVA table. Comment on what this output tells you about the GPA across college stduents. Provide formal hypotheses, p-values, and conclusions.

```{r}
amodA = aov(GPA~LarkOwl+Stress, data=Sleepstudy)
summary(amodA)
```

Hypotheses:

H0: μ1 = μ2 = μ3 (1= Lark, 2 = Owl, 3= Neither) (Mean GPA across LarkOwl status are all same)
Ha: Some μi != μk (There is at least one mean GPA of one LarkOwl status that is different from the mean GPA of another LarkOwl status)

P-value: 0.262

Conclusion:
Because p-value is not small enough (0.262), we fail to reject the null hypothesis. Mean GPA across LarkOwl statuses are all same.

Hypotheses:

H0: α1 = α2 = 0 (1 = Normal, 2 = High) (The affect for both stress scores are zeros)
Ha: One of αi != 0 (There is at least one stress score that does not have zero for the affect)

p-value: 0.012

Conclusion: Because our p-value is small enough (0.012), we reject the null hypothesis. There is at least one stress score that does not have zero for the affect. 


2)  Construct residual plots and comment on the conditions of equality of variances and normality of residuals for the model that you created. 

```{r}
amodA
plot(amodA, 1:2)
tapply(Sleepstudy$GPA, Sleepstudy$LarkOwl, sd)
tapply(Sleepstudy$GPA, Sleepstudy$Stress, sd)
```

Variances: In general, we have constant variances, but as fitted values increase, the variances (vertical distances) are decreasing. In addition, the standard deviations for different combinations of LarkOwl and Stress predictors are roughly similar. 

Normality: The normal Q-Q line shows some skewness on the left tail. 


