---
title: "Methods for Choosing Predictors"
author: "Alyssa Warnock, Helen Johnston, Ashley Wade, Hunter Barbee"
date: "11/28/2021"
output:
  word_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Consider the SleepStudy file described on Sakai.  In this exercise, you will consider models to predict Grade Point Average (GPA) using any of the predictors.  
```{r}
sleep <- read.csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/SleepStudy.csv")

source("https://raw.githubusercontent.com/JA-McLean/STOR455/master/scripts/ShowSubsets.R")

Full=lm(GPA~factor(Gender)+factor(ClassYear)+factor(EarlyClass) + LarkOwl + NumEarlyClass + ClassesMissed + CognitionZscore + PoorSleepQuality + DepressionScore + AnxietyScore + StressScore + DepressionStatus + AnxietyStatus + Stress + DASScore + Happiness + AlcoholUse + Drinks + WeekdayBed + WeekdayRise + WeekdaySleep + WeekendBed + WeekendRise + WeekendSleep + AverageSleep + AllNighter, data=sleep)
```


Use forward selection until you have a  model for GPA.  
```{r}
none=lm(GPA~1,data=sleep)

MSE=(summary(Full)$sigma)^2
   #Specify the direction
step(none,scope=list(upper=Full),scale=MSE, direction= "forward", trace = FALSE)
```

Use backwards elimination until you have a model for GPA.   
```{r}
Full=lm(GPA~factor(Gender)+factor(ClassYear)+factor(EarlyClass) + LarkOwl + NumEarlyClass + ClassesMissed + CognitionZscore + PoorSleepQuality + DepressionScore + AnxietyScore + StressScore + DepressionStatus + AnxietyStatus + Stress + DASScore + Happiness + AlcoholUse + Drinks + WeekdayBed + WeekdayRise + WeekdaySleep + WeekendBed + WeekendRise + WeekendSleep + AverageSleep + AllNighter, data=sleep)
# Find the MSE for the full model

MSE=(summary(Full)$sigma)^2
# Backward: use the step( ) command starting with the full model
#MSE = variance of the residuals 

step(Full,scale=MSE, trace = FALSE)
```

Use a “best subsets” procedure to determine which predictors would explain the most variability in GPA.  

```{r}
source("https://raw.githubusercontent.com/JA-McLean/STOR455/master/scripts/ShowSubsets.R")
library(leaps)
all=regsubsets(GPA~factor(Gender)+factor(ClassYear)+factor(EarlyClass) + LarkOwl + NumEarlyClass + ClassesMissed + CognitionZscore + PoorSleepQuality + DepressionScore + AnxietyScore + StressScore + DepressionStatus + AnxietyStatus + Stress + DASScore + Happiness + AlcoholUse + Drinks + WeekdayBed + WeekdayRise + WeekdaySleep + WeekendBed + WeekendRise + WeekendSleep + AverageSleep + AllNighter, data= sleep, nvmax=26)
ShowSubsets(all)
```
Predictors: factor(ClassYear), ClassesMissed, CognitionZscore,  PoorSleepQuality, StressScore, DASScore, Drinks, WeekdaySleep, WeekendRise, AllNighter.

